Accelerating adaptive ultrasound algorithms using graphics processing units \vspace{0.5cm}

### 2014-03-22 ThesisJP/Background/background.tex
Started on section about ultrasound field simulations.

If the width of a piezoelectric element is less than the wavelength of the transmitted signal the element is said to be \textit{omnidirectional}, hence energy is distributed or sensed equally in all directions. When multiple omnidirectional elements are arranged in a linear array the transmitted pulse will get shaped into a beam\footnote{A beam refers to the pulse in space integrated across time.} of ultrasound by diffraction as the size of the array  (the \textit{aperture}) grows. The 3 dB angular width of this beam, and also the angular resolution of the array, is given by \cite{AngelUltrasound}:
\begin{align}\label{eq:res}
\delta\theta_a = \frac{\lambda}{D} = \frac{c}{fD},
\end{align}
\nomi{$\delta\theta_a$}{Angular resolution of an aperture}where \nom{$\lambda$}{Wavelength} and \nom{$f$}{Frequency} is the ultrasound wavelength and frequency respectively, $c$ is the speed of sound, and \nom{$D$}{Aperture size} is the aperture width. With an average speed-of-sound in human tissue of $1540\frac{\text{m}}{\text{s}}$, the angular resolution or opening angle of a 2 cm cardiac probe operating at 3 MHz is around 1.5\degree, which at 8 cm range gives a lateral resolution of around 2 mm. The axial resolution is inverse proportional to the pulse length, and is typically better than the lateral resolution.

@@

When generating a B-mode image we have to insure that the spatial and temporal sampling are sufficient. We have already seen that the lateral resolution of an ultrasound probe is given by (\ref{eq:res}).  Hergum \textit{et al.} \cite{Hergum2009} have further shown that the lateral frequency support for a system which both emit and receive energy is governed by the sum of the transmit (\nom{tx}{Transmit}) and receive (\nom{rx}{Receive}) aperture. The lateral resolution of what is known as an active imaging system is found to be
\begin{align}
\delta\theta = \frac{\lambda}{D_{tx} + D_{rx}}.
\end{align}
\nomi{$\delta\theta$}{Angular resolution of a two-way imaging system}\nomi{$D_{tx}$}{Aperture size on transmit}\nomi{$D_{rx}$}{Aperture size on receive}Then, in order to properly sample an imaging sector of $\theta$ radians we have to acquire 
\begin{align}
N_\theta = \frac{\theta}{\delta\theta}
\end{align}
\nomi{$N_\theta$}{Number of lines required for proper sampling of an imaging sector}A-mode images uniformly distributed across the sector. If \nom{$r$}{Imaging range.} is the maximum range we want to image we get an imaging frame rate of
\begin{align}
f_{\text{FPS}} = \frac{c}{2rN_\theta}
\end{align} 
frames per second (\nom{FPS}{Frames per second}). With our 2 cm cardiac probe we get a frame rate of around $55$ FPS when imaging a 70\degree{} sector with $r=15$ cm.

@@

\section{Field simulations}\label{sec:field}

Small chapter about different simulation tools (See hos paper).

>>>>

\subsection{Beamforming}
If the width of a piezoelectric element is less than the wavelength of the transmitted signal the element is said to be \textit{omnidirectional}, hence energy is distributed or sensed equally in all directions. When multiple omnidirectional elements are arranged in a linear array the transmitted pulse will get shaped into a beam\footnote{A beam refers to the pulse in space integrated across time.} of ultrasound by diffraction as the size of the array  (the \textit{aperture}) grows. The 3 dB angular width of this beam, and also the angular resolution of the array, is given by \cite{AngelUltrasound}:
\begin{align}\label{eq:res}
\delta\theta_a = \frac{\lambda}{D} = \frac{c}{fD},
\end{align}
\nomi{$\delta\theta_a$}{Angular resolution of an aperture}where \nom{$\lambda$}{Wavelength} and \nom{$f$}{Frequency} is the ultrasound wavelength and frequency respectively, $c$ is the speed of sound, and \nom{$D$}{Aperture size} is the aperture width. With an average speed-of-sound in human tissue of $1540\frac{\text{m}}{\text{s}}$, the angular resolution or opening angle of a 2 cm cardiac probe operating at 3 MHz is around 1.5\degree, which at 8 cm range gives a lateral resolution of around 2 mm. The axial resolution is inverse proportional to the pulse length, and is typically better than the lateral resolution.

@@

\subsection{Sampling and processing complexity}
When generating a B-mode image we have to insure that the spatial and temporal sampling are sufficient. We have already seen that the lateral resolution of an ultrasound probe is given by (\ref{eq:res}).  Hergum \textit{et al.} \cite{Hergum2009} have further shown that the lateral frequency support for a system which both emit and receive energy is governed by the sum of the transmit (\nom{tx}{Transmit}) and receive (\nom{rx}{Receive}) aperture. The lateral resolution of what is known as an active imaging system is found to be
\begin{align}
\delta\theta = \frac{\lambda}{D_{tx} + D_{rx}}.
\end{align}
\nomi{$\delta\theta$}{Angular resolution of a two-way imaging system}\nomi{$D_{tx}$}{Aperture size on transmit}\nomi{$D_{rx}$}{Aperture size on receive}Then, in order to properly sample an imaging sector of $\theta$ radians we have to acquire 
\begin{align}
N_\theta = \frac{\theta}{\delta\theta}
\end{align}
\nomi{$N_\theta$}{Number of lines required for proper sampling of an imaging sector}A-mode images uniformly distributed across the sector. If \nom{$r$}{Imaging range.} is the maximum range we want to image we get an imaging frame rate of
\begin{align}
f_{\text{FPS}} = \frac{c}{2rN_\theta}
\end{align} 
frames per second (\nom{FPS}{Frames per second}). With our 2 cm cardiac probe we get a frame rate of around $55$ FPS when imaging a 70\degree{} sector with $r=15$ cm.

@@

\section{Ultrasound field simulations}\label{sec:field}
Simulation of ultrasound probes are heavily used by researchers in order to test out new algorithms and probe designs. A  typical simulator for ultrasound probes works by approximating the probe surface with discrete points or patches. In software it is then possible to transmit a pulse from a given probe and measure its response in space and time. The most popular simulator for ultrasound field simulations is Field II \cite{Jensen1992}. All simulations in both \textbf{Paper II} and \textbf{III} are produced with this simulator.

Field II. Spatial impulse response.

Ultrasim. Fusk. Propose. Explain the principle behind all.

What is the computational complexity of Filed II and Ultrasim.

Simulators implemented on the GPU.
